\thetitle

### 2017-07-03 compared.tex
convert terminology

\section{Context algebras compared to Voevodsky's B-systems}\label{sec:esys_compared}
In an effort to formalize the simplicial set model of Martin-L\"of type theory
with the univalence axiom, Voevodsky has studied various notions of models of
type theory, taking Cartmell's \emph{contextual categories} as a starting
point. Voevodsky's definition of a \emph{C-system} is equivalent to that of Cartmell
\cite{VV_C-systems_quotients}. There are several techniques of defining 
C-systems: in \cite{VV_Csys_univ}, C-systems are constructed out of a universe
in a category; in \cite{VV_C-systems_quotients}, sub-C-systems and regular
quotients of C-systems are defined; in \cite{VV_C-systems_monad}, a C-system
is constructed out of a monad $R$ on the category of sets and a left $R$-module
with values in $\mathbf{Set}$.

@@

\subsection{Stratified context algebras}

In this subsection we define the notion of stratified context algebras. Then we show that the category $\mathbf{Tel}_s$ of stratified context algebras with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Tel}$.

@@

\begin{defn}
A pre-Context algebras is said to be \define{stratified} if its underlying category is
stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
stratified pre-Context algebras with stratified E-homomorphisms between them is denoted by
$\mathbf{Tel}_s$.
\end{defn}

\begin{cor}\label{lem:strat_full}
The category of stratified pre-Context algebras is a full subcategory of the category of
pre-Context algebras.
\end{cor}

\subsection{The embedding of B0-systems into pre-Context algebras}

Voevodsky distinguishes between pre-B-systems, which is an essentially algebraic theory with
only operations and no equalities; B0-systems, which is an extension of the
essentially algebraic theory of pre-B-systems which ensures that all the operations behave
well with two sets of the operations ($\eft$ and $\ebd$); and B-systems, which extends the 
theory of B0-systems with equalities ensuring that the rest of the operations 
are compatible with each other. Voevodsky, furthermore distinguishes
non-unital and unital versions of these, and noted that not all B-homomorphisms
between unital B-systems are unital. The same distinguishment could be made for 
Context algebras by forgetting the projection structure, but we have not been able to construct a
functor from unital B0-systems to unital pre-Context algebras. Therefore, we shall only
consider non-unital B0-systems and unital B-systems. In this section we will
recall the definition of non-unital B0-systems and construct a full and faithful
functor from them to the category of pre-Context algebras.

One reason for introducing B-systems by first introducing pre-B-systems is that
it is clear in the definition of pre-B-systems that there are only countably
many operations. Since we are interested in embedding the B-systems into context algebras,
we shall organize the definition of B-systems a bit differently, but of course
the resulting definition of B0-systems and B-systems will be (trivially) 
equivalent.

@@

Now that we have defined the category of B0systems, we can start defining the
embedding into pre-Context algebras.

@@

\begin{defn}
Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
of a pre-Context algebras. We write $\mathcal{E}(\mathbb{B})$ for the resulting
pre-Context algebras. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Tel}_s$.
\end{defn}

@@

\subsection{B-systems are stratified context algebras}
There are two properties which hold for pre-Context algebras of the form $\mathcal{E}
(\mathbb{B})$ for some B0-system $\mathbb{B}$, 
which stand in the way of proving that the category of stratified
pre-Context algebras is equivalent to the category of B0-systems. The first is that
for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
we have an isomorphism $T(\ctxext{A}{P})\cong\bigsqcup_{x\in T(A)}T(S_x(P))$.
The second, which depends on the first, is that for any $(x,u)\in T(\ctxext{A}{P})$
we have $S(x,u)\jdeq S_u\circ S_x/P$. In other words, both the term structure
and the pre-substitution structure of $\mathcal{E}(\mathbb{B})$ are compatible
with the categorical structure of the underlying category $U(\mathbb{B})$.
However, to prove the corresponding facts for context algebras, we needed the full
set of conditions on unital context algebras: to prove that the term structure of
$U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
that the operations $S_x$ are compatible with the pairing function we needed
that they are E-homomorphisms.

Presumably, we can assume these compatibility conditions on pre-Context algebras in order
to prove the desired equivalence of B0-systems and stratified pre-Context algebras. However,
that seems somewhat unnatural, especially given that they automatically hold
for context algebras. Thus, for the remainder of this section we set out to show that
the category of unital B-systems is equivalent to the category of stratified 
unital context algebras.

@@

\begin{lem}
For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
stratified non-unital context algebras. 
\end{lem}

@@

\begin{defn}
For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
projection structure. We thus obtain a functor 
$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ from unital B-systems to context algebras.
\end{defn}

\begin{constr}
We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
from non-unital B0-systems to non-unital pre-Context algebras, to a functor from
unital B-systems to context algebras.

We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
pre-Context algebras. We construct by induction on $m\in \mathbb{N}$, a term of type 
$T(m,W_{(m,X)}(X))$ with the property that that 
\begin{equation*}
S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
\end{equation*} 
for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
condition is part of the inductive construction, because we need it on the way.

@@

\begin{thm}
The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Tel}$ lifts to an equivalence
to the subcategory $\mathbf{Tel}_s$ of stratified context algebras: i.e.~we have a
commuting diagram
\begin{equation*}
\begin{tikzcd}
& \mathbf{Tel}_s \arrow[dr,"\subseteq"]\\
\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Tel}
\end{tikzcd}
\end{equation*}
\end{thm}

\begin{proof}
We already have shown that $\mathcal{E}$ factors through the stratified
Context algebras and that it is full and faithful, so it remains
to show that every stratified context algebras is in the image of $\mathcal{E}$. In
other words, for every stratified context algebras $\mathbb{E}$ we need to construct
a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.

Let $\mathbb{E}$ be a stratified telescopi type system. From the stratification functor
$L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
presheaf $B(\mathbb{E}):\op{(\mathbb{N},\leq)}\to\mathbf{Set}$ given by
\begin{align*}
B(\mathbb{E})_n & \defeq \{X\in\mathrm{Ob}(\mathbb{C}_{\mathbb{E}})\mid L(X)\jdeq n\}
\end{align*}
and $\eft(X)\defeq\mathrm{cod}(p_X)$. We define the term structure of $B(\mathbb{E})_n$
to be given by
\begin{equation*}
\tilde{B}(\mathbb{E})_n \defeq \bigsqcup_{X\in B(\mathbb{E})_n} T(p_X)
\end{equation*}
with the projection map $\ebd[n]\defeq\pi_1:\tilde{B}(\mathbb{E})_n\to B(\mathbb{E})_n$.
This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
$B:\mathbf{Tel}_s\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
$U(B(\mathbb{E}))\cong\cat{F}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.

The weakening and substitution structures on
$B(\mathbb{E})$ are given by $W_X\defeq B(W_{p_X})$ and $S_{(X,x)}\defeq B(S_x)$,
respectively. Since these functors are assumed to be stratified, they define
homomorphisms of B-frameworks. Because a weakening system of a context algebras is
required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
$W_{\catid{\Gamma}}\jdeq\catid{\cat{F}_{\mathbb{E}}/\Gamma}$, it follows
immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
that of $\mathbb{E}$. 
\end{proof}

@@

\section{Context algebras compared to comprehension categories and to categories with families}
\subsection{Telescopic categories with families}

@@

There are of course many more sensible examples of categories with families,
but the following class of examples causes us to doubt whether it is possible
at all to have a faithful functor from the category of categories
with families to the category of context algebras.

@@

\begin{rmk}
Although a terminal object is usually assumed to exist, not every object 
of a category with families is assumed to be `fibrant', i.e.~it is
not the case that every $X\to 1$ is a composition of projections $p_A$.
We could only hope that there is a functor from D-sytems to context algebras without
terminal objects.
\end{rmk}

@@

The following two definitions define subclasses of telescopic D-systems and 
coherent telescopic D-systems, and the latter structure is the same as an E'-system
(without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
to obtain a telescopic D-system from any comprehension category. This construction
will also generate a telescopic D-system out of any D-system, because D-systems
are comprehension categories.

@@

\begin{thm}
An E'-system without empty context is the same thing as a coherent telescopic D-system.
\end{thm}

\begin{proof}
Let $\mathbb{E}$ be an E'-system without empty context. Then we define the
presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
\begin{equation*}
\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
\end{equation*}
and on morphisms by the action on objects of $f^\ast$, for each $f:\Delta\to
\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of E'-systems, it
follows that this defines a contravariant functor.

We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
$\cat{C}$, and by sending $f:(\Delta,f^\ast(A))\to (\Gamma,A)$ to the pullback
square
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxext{\Delta}{f^\ast(A)} \arrow[d,swap,"{f^\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
This is also functorial because of conditions (b) and (c) in the definition of
E'-systems. Thus, we have defined a D-system
\begin{equation*}
\begin{tikzcd}
\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
\end{tikzcd}
\end{equation*}
and it is straightforward to verify that this D-system is telescopic and coherent.

@@

\subsection{Coherent comprehension categories}
In this section we recall Jacobs' notion of comprehension categories, and 
Dybjer's notion of categories with families. It is shown in 
\cite{Hofmann_syntax_semantics} that the category of categoriew with families
is equivalent to the category of categories with attributes, which forms a
subcategory of comprehension categories. Thus, for our purpose of comparing
Context algebras with various existing notions of models of type dependency, it suffice
to make the comparison with comprehension categories.

In this section we do the preparateions for this comparison. After having
introduced categories with families, we prove that categories with families
with the extra structure of `being closed under family extension' in a coherent
way, are E'-systems without empty context.

@@

Each comprehension category determines a subcategory $\cat{F}$ generated by
the projections, as we shall see soon. However, there is no reason why
this category should possess a terminal object, so we naturally have to retreat
to E'-systems without empty contex. Even when we make this generalization, 
the construction of a canonical functor in which families are represented by 
compositions of projections is still not possible without a further coherence 
condition. To make sure that
the pullback operations of a comprehension category extend to the category
$\cat{F}$, we introduce \emph{coherent} comprehension categories. So we only have a functor $U$
from the category $\mathbf{Dsys}$ of coherent D-systems to the category
$\mathbf{E'sys_{\circ}}$ of E'-systems without empty context.
We see that it is not evident at all to turn categories with families into
Context algebras, even if we ignore the issue that categories with families map
naturally into E'-systems without empty context.
Thus it seems that the type theories of categories with families and of context algebras
are not directly comparable.

We end this section by showing that the functor 
$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ restricts to a full and faithful 
functor on the coherent D-systems which satisfy the condition that every 
composition of projections is a composition of projections in precisely one way.
We call coherent D-systems with this property
locally well-founded. Among the locally well-founded D-systems are the B-systems.
As a by-product, we get that the full and faithful functor from B-systems to
Context algebras factors through D-systems.

To get from comprehension categories to E'-systems without empty contexts, we
need to turn comprehension categories into telescopic D-systems.

@@

\begin{thm}
Let $\mathcal{P}$ be a comprehension category. Then the following are equivalent:
\begin{enumerate}
\item The choice of functors $\Lambda(f)$ and lifts $q(f,A)$ can be extended to the 
structure of an E'-system without empty context.
\item For any
$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$
and for any $f:\Delta\to\Gamma$ in $\cat{C}$, the equalities
$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$ and
$\pi_2(f,A)\jdeq\pi_2(f,A')$ hold. 
\end{enumerate}
\end{thm}

\begin{proof}
Suppose we can extend $\Lambda(f)$ for all $f:\Delta\to\Gamma$, to form commuting
diagrams
\begin{equation*}
\begin{tikzcd}
\Lambda(\Gamma) \arrow[d,swap,"\mathcal{P}"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"\mathcal{P}"] \\
\cat{F}/\Gamma \arrow[r,densely dotted,"{f^\ast}"] & \cat{F}/\Delta
\end{tikzcd}
\end{equation*}
of functors, forming an E'-system without empty context, where $\mathcal{P}(A)
\defeq \mathcal{P}(A_1)\circ\cdots\circ\mathcal{P}(A_n)$ for any
$A\jdeq (A_1,\ldots,A_n)\in\Lambda(\Gamma)$. Then it must hold that for any
$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$, we have
$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$. Also, for such
$A$ and $A'$ it follows from the way $\Lambda(f)$ is defined, 
that $\Lambda(\pi_2(f,A))\jdeq \Lambda(f)/A\jdeq
\Lambda(f)/A' \jdeq \Lambda(\pi_2(f,A'))$, and therefore we also get the equalities
\begin{equation*}
\pi_2(f,A)\jdeq \pi_2(\pi_2(f,A),())\jdeq \pi_2(\pi_2(f,A'),()) \jdeq \pi_2(f,A'),
\end{equation*}
so the desired equality holds.

@@

\begin{conj}
The functor from coherent split comprehension categories to E'-systems without
empty context is full and faithful.
\end{conj}

\begin{proof}
Let $\mathcal{E}$ be an E'-system. Define $\cat{E}$ to be the full subcategory
of $\cat{C}^\to$ whose objects are the morphisms of $\cat{F}$. Thus, for any
$\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
morphisms the context algebras $\mathcal{E}/\Gamma$.

@@

\begin{thm}
The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ to the
locally well-founded D-systems is full and faithful.
\end{thm}

@@

\begin{thm}
The image of $U$ is the subcategory of locally stratified E'-systems without
empty context. 
\end{thm}

\begin{proof}
For any $\Gamma\in\mathbb{D}$, there is an obvious stratification functor
$\Lambda(\Gamma)\to (\N,\geq)$, so $U(\mathbb{D})$ is a locally stratified
E'-system.

Now let $\mathbb{E}$ be a locally stratified E'-system without empty context.
For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of
families of length one. Since any $f^\ast$ is stratified, it follows that
$\mathrm{Ty}$ is a presheaf on $\cat{C}$. The choice of pullback squares
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxext{\Delta}{f^\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,swap,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
is functorial by assumption, so we have defined a category with families. It
is straigtforward to verify that this category with families is locally well-founded,
and that its construction is inverse to $U$. 
\end{proof}

>>>>

\section{E-systems compared to Voevodsky's B-systems}\label{sec:esys_compared}
In an effort to formalize the simplicial set model of Martin-L\"of type theory
with the univalence axiom, Voevodsky has studied various notions of models of
type theory, taking Cartmell's \emph{contextual categories} as a starting
point. Voevodsky's definition of a \emph{C-system} is equivalent to that of Cartmell
\cite{VV_C-systems_quotients}. There are several techniques of defining 
C-systems: in \cite{VV_Csys_univ}, C-systems are constructed out of a universe
in a category; in \cite{VV_C-systems_quotients}, sub-C-systems and regular
quotients of C-systems are defined; in \cite{VV_C-systems_monad}, a C-system
is constructed out of a monad $R$ on the category of sets and a left $R$-module
with values in $\mathbf{Set}$.

@@

\subsection{Stratified E-systems}

In this subsection we define the notion of stratified E-systems. Then we show that the category $\mathbf{Esys}_s$ of stratified E-systems with homomorphisms preserving the stratification, is a full subcategory of $\mathbf{Esys}$.

@@

\begin{defn}
A pre-E-systems is said to be \define{stratified} if its underlying category is
stratified and if each $W_A$ and $S_x$ is a stratified functor. The category of
stratified pre-E-systems with stratified E-homomorphisms between them is denoted by
$\mathbf{Esys}_s$.
\end{defn}

\begin{cor}\label{lem:strat_full}
The category of stratified pre-E-systems is a full subcategory of the category of
pre-E-systems.
\end{cor}

\subsection{The embedding of B0-systems into pre-E-systems}

Voevodsky distinguishes between pre-B-systems, which is an essentially algebraic theory with
only operations and no equalities; B0-systems, which is an extension of the
essentially algebraic theory of pre-B-systems which ensures that all the operations behave
well with two sets of the operations ($\eft$ and $\ebd$); and B-systems, which extends the 
theory of B0-systems with equalities ensuring that the rest of the operations 
are compatible with each other. Voevodsky, furthermore distinguishes
non-unital and unital versions of these, and noted that not all B-homomorphisms
between unital B-systems are unital. The same distinguishment could be made for 
E-systems by forgetting the projection structure, but we have not been able to construct a
functor from unital B0-systems to unital pre-E-systems. Therefore, we shall only
consider non-unital B0-systems and unital B-systems. In this section we will
recall the definition of non-unital B0-systems and construct a full and faithful
functor from them to the category of pre-E-systems.

One reason for introducing B-systems by first introducing pre-B-systems is that
it is clear in the definition of pre-B-systems that there are only countably
many operations. Since we are interested in embedding the B-systems into E-systems,
we shall organize the definition of B-systems a bit differently, but of course
the resulting definition of B0-systems and B-systems will be (trivially) 
equivalent.

@@

Now that we have defined the category of B0systems, we can start defining the
embedding into pre-E-systems.

@@

\begin{defn}
Let $\mathbb{B}$ be a B0-system. Then $U(\mathbb{B})$ can be given the structure
of a pre-E-systems. We write $\mathcal{E}(\mathbb{B})$ for the resulting
pre-E-systems. Thus we get a functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys}_s$.
\end{defn}

@@

\subsection{B-systems are stratified E-systems}
There are two properties which hold for pre-E-systems of the form $\mathcal{E}
(\mathbb{B})$ for some B0-system $\mathbb{B}$, 
which stand in the way of proving that the category of stratified
pre-E-systems is equivalent to the category of B0-systems. The first is that
for any $A\in U(\mathbb{B})/\Gamma$ and $P\in U(\mathbb{B})/\ctxext{\Gamma}{A}$,
we have an isomorphism $T(\ctxext{A}{P})\cong\bigsqcup_{x\in T(A)}T(S_x(P))$.
The second, which depends on the first, is that for any $(x,u)\in T(\ctxext{A}{P})$
we have $S(x,u)\jdeq S_u\circ S_x/P$. In other words, both the term structure
and the pre-substitution structure of $\mathcal{E}(\mathbb{B})$ are compatible
with the categorical structure of the underlying category $U(\mathbb{B})$.
However, to prove the corresponding facts for E-systems, we needed the full
set of conditions on unital E-systems: to prove that the term structure of
$U(\mathbb{B})$ is compatible with composition we needed units, and to prove 
that the operations $S_x$ are compatible with the pairing function we needed
that they are E-homomorphisms.

Presumably, we can assume these compatibility conditions on pre-E-systems in order
to prove the desired equivalence of B0-systems and stratified pre-E-systems. However,
that seems somewhat unnatural, especially given that they automatically hold
for E-systems. Thus, for the remainder of this section we set out to show that
the category of unital B-systems is equivalent to the category of stratified 
unital E-systems.

@@

\begin{lem}
For any non-unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ is a 
stratified non-unital E-systems. 
\end{lem}

@@

\begin{defn}
For any unital B-system $\mathbb{B}$, $\mathcal{E}(\mathbb{B})$ has a
projection structure. We thus obtain a functor 
$\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ from unital B-systems to E-systems.
\end{defn}

\begin{constr}
We first need to extend the functor $\mathcal{E}:\mathbf{B0sys}\to\mathbf{Esys_p}$
from non-unital B0-systems to non-unital pre-E-systems, to a functor from
unital B-systems to E-systems.

We need to give $\mathcal{E}(\mathbb{B})$ the structure of a unital
pre-E-systems. We construct by induction on $m\in \mathbb{N}$, a term of type 
$T(m,W_{(m,X)}(X))$ with the property that that 
\begin{equation*}
S_{\tfid{(m,X)}}\circ W_{(m,X)}/X\jdeq\catid{U(\mathbb{B})/X},
\end{equation*} 
for every $(n+m,X)\leq (n,\Gamma)$ by induction on $m$. The extra
condition is part of the inductive construction, because we need it on the way.

@@

\begin{thm}
The functor $\mathcal{E}:\mathbf{Bsys}\to\mathbf{Esys}$ lifts to an equivalence
to the subcategory $\mathbf{Esys}_s$ of stratified E-systems: i.e.~we have a
commuting diagram
\begin{equation*}
\begin{tikzcd}
& \mathbf{Esys}_s \arrow[dr,"\subseteq"]\\
\mathbf{Bsys} \arrow[ur,"\simeq"] \arrow[rr,swap,"\mathcal{E}"] & & \mathbf{Esys}
\end{tikzcd}
\end{equation*}
\end{thm}

\begin{proof}
We already have shown that $\mathcal{E}$ factors through the stratified
E-systems and that it is full and faithful, so it remains
to show that every stratified E-systems is in the image of $\mathcal{E}$. In
other words, for every stratified E-systems $\mathbb{E}$ we need to construct
a B0-system $\mathbb{B}$ such that $\mathcal{E}(\mathbb{B})\cong\mathbb{E}$.

Let $\mathbb{E}$ be a stratified telescopi type system. From the stratification functor
$L:\mathbb{C}_{\mathbb{E}}\to (\mathbb{N},\geq)$ we obtain the pre-image
presheaf $B(\mathbb{E}):\op{(\mathbb{N},\leq)}\to\mathbf{Set}$ given by
\begin{align*}
B(\mathbb{E})_n & \defeq \{X\in\mathrm{Ob}(\mathbb{C}_{\mathbb{E}})\mid L(X)\jdeq n\}
\end{align*}
and $\eft(X)\defeq\mathrm{cod}(p_X)$. We define the term structure of $B(\mathbb{E})_n$
to be given by
\begin{equation*}
\tilde{B}(\mathbb{E})_n \defeq \bigsqcup_{X\in B(\mathbb{E})_n} T(p_X)
\end{equation*}
with the projection map $\ebd[n]\defeq\pi_1:\tilde{B}(\mathbb{E})_n\to B(\mathbb{E})_n$.
This defines the B-framework $B(\mathbb{E})$, and we obtain a functor
$B:\mathbf{Esys}_s\to\mathbf{B0sys}$. It is immediate that we get isomorphisms
$U(B(\mathbb{E}))\cong\cat{F}_{\mathbb{E}}$, natural in $\mathbb{E}$. Also, we
get natural isomorphisms $B(\mathbb{E}/\Gamma)\cong B(\mathbb{E})/\Gamma$.

The weakening and substitution structures on
$B(\mathbb{E})$ are given by $W_X\defeq B(W_{p_X})$ and $S_{(X,x)}\defeq B(S_x)$,
respectively. Since these functors are assumed to be stratified, they define
homomorphisms of B-frameworks. Because a weakening system of a E-systems is
required to be functorial, i.e.~$W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ and
$W_{\catid{\Gamma}}\jdeq\catid{\cat{F}_{\mathbb{E}}/\Gamma}$, it follows
immediately that the weakening structure on $U(B(\mathbb{E}))$ is isomorphic to
that of $\mathbb{E}$. 
\end{proof}

@@

\section{E-systems compared to comprehension categories and to categories with families}
\subsection{Telescopic categories with families}

@@

There are of course many more sensible examples of categories with families,
but the following class of examples causes us to doubt whether it is possible
at all to have a faithful functor from the category of categories
with families to the category of E-systems.

@@

\begin{rmk}
Although a terminal object is usually assumed to exist, not every object 
of a category with families is assumed to be `fibrant', i.e.~it is
not the case that every $X\to 1$ is a composition of projections $p_A$.
We could only hope that there is a functor from D-sytems to E-systems without
terminal objects.
\end{rmk}

@@

The following two definitions define subclasses of telescopic D-systems and 
coherent telescopic D-systems, and the latter structure is the same as an CE-system
(without empty context). In \autoref{defn:comp_to_telescopic}, we will show how
to obtain a telescopic D-system from any comprehension category. This construction
will also generate a telescopic D-system out of any D-system, because D-systems
are comprehension categories.

@@

\begin{thm}
An CE-system without empty context is the same thing as a coherent telescopic D-system.
\end{thm}

\begin{proof}
Let $\mathbb{E}$ be an CE-system without empty context. Then we define the
presheaf $\mathrm{Fam}$ on $\cat{C}$ on objects by
\begin{equation*}
\mathrm{Fam}(\Gamma)\defeq \mathrm{Obj}(\cat{F}/\Gamma)
\end{equation*}
and on morphisms by the action on objects of $f^\ast$, for each $f:\Delta\to
\Gamma$ in $\cat{C}$. By condition (b) and (c) in the definition of CE-systems, it
follows that this defines a contravariant functor.

We define a functor $\mathcal{P}:\int_{\cat{C}}\to\cat{C}^\square$ by sending
the pair $(\Gamma,A)$ to the morphism $A:\ctxext{\Gamma}{A}\to\Gamma$ in
$\cat{C}$, and by sending $f:(\Delta,f^\ast(A))\to (\Gamma,A)$ to the pullback
square
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxext{\Delta}{f^\ast(A)} \arrow[d,swap,"{f^\ast(A)}"] \arrow[r,"{\pi_2(f,A)}"] &
\ctxext{\Gamma}{A} \arrow[d,"A"] \\ \Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
This is also functorial because of conditions (b) and (c) in the definition of
CE-systems. Thus, we have defined a D-system
\begin{equation*}
\begin{tikzcd}
\int_{\cat{C}}\mathrm{Fam} \arrow[rr,"{\mathcal{P}}"] \arrow[dr,"{\pi}"] & & \cat{C}^\square \arrow[dl,"{\mathrm{cod}}"] \\ & \cat{C}
\end{tikzcd}
\end{equation*}
and it is straightforward to verify that this D-system is telescopic and coherent.

@@

\subsection{Coherent comprehension categories}
In this section we recall Jacobs' notion of comprehension categories, and 
Dybjer's notion of categories with families. It is shown in 
\cite{Hofmann_syntax_semantics} that the category of categoriew with families
is equivalent to the category of categories with attributes, which forms a
subcategory of comprehension categories. Thus, for our purpose of comparing
E-systems with various existing notions of models of type dependency, it suffice
to make the comparison with comprehension categories.

In this section we do the preparateions for this comparison. After having
introduced categories with families, we prove that categories with families
with the extra structure of `being closed under family extension' in a coherent
way, are CE-systems without empty context.

@@

Each comprehension category determines a subcategory $\cat{F}$ generated by
the projections, as we shall see soon. However, there is no reason why
this category should possess a terminal object, so we naturally have to retreat
to CE-systems without empty contex. Even when we make this generalization, 
the construction of a canonical functor in which families are represented by 
compositions of projections is still not possible without a further coherence 
condition. To make sure that
the pullback operations of a comprehension category extend to the category
$\cat{F}$, we introduce \emph{coherent} comprehension categories. So we only have a functor $U$
from the category $\mathbf{Dsys}$ of coherent D-systems to the category
$\mathbf{CEsys_{\circ}}$ of CE-systems without empty context.
We see that it is not evident at all to turn categories with families into
E-systems, even if we ignore the issue that categories with families map
naturally into CE-systems without empty context.
Thus it seems that the type theories of categories with families and of E-systems
are not directly comparable.

We end this section by showing that the functor 
$U:\mathbf{Dsys}\to\mathbf{CEsys_{\circ}}$ restricts to a full and faithful 
functor on the coherent D-systems which satisfy the condition that every 
composition of projections is a composition of projections in precisely one way.
We call coherent D-systems with this property
locally well-founded. Among the locally well-founded D-systems are the B-systems.
As a by-product, we get that the full and faithful functor from B-systems to
E-systems factors through D-systems.

To get from comprehension categories to CE-systems without empty contexts, we
need to turn comprehension categories into telescopic D-systems.

@@

\begin{thm}
Let $\mathcal{P}$ be a comprehension category. Then the following are equivalent:
\begin{enumerate}
\item The choice of functors $\Lambda(f)$ and lifts $q(f,A)$ can be extended to the 
structure of an CE-system without empty context.
\item For any
$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$
and for any $f:\Delta\to\Gamma$ in $\cat{C}$, the equalities
$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$ and
$\pi_2(f,A)\jdeq\pi_2(f,A')$ hold. 
\end{enumerate}
\end{thm}

\begin{proof}
Suppose we can extend $\Lambda(f)$ for all $f:\Delta\to\Gamma$, to form commuting
diagrams
\begin{equation*}
\begin{tikzcd}
\Lambda(\Gamma) \arrow[d,swap,"\mathcal{P}"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"\mathcal{P}"] \\
\cat{F}/\Gamma \arrow[r,densely dotted,"{f^\ast}"] & \cat{F}/\Delta
\end{tikzcd}
\end{equation*}
of functors, forming an CE-system without empty context, where $\mathcal{P}(A)
\defeq \mathcal{P}(A_1)\circ\cdots\circ\mathcal{P}(A_n)$ for any
$A\jdeq (A_1,\ldots,A_n)\in\Lambda(\Gamma)$. Then it must hold that for any
$A,A'\in\Lambda(\Gamma)$ such that $\mathcal{P}(A)\jdeq\mathcal{P}(A')$, we have
$\mathcal{P}(\Lambda(f,A))\jdeq\mathcal{P}(\Lambda(f,A'))$. Also, for such
$A$ and $A'$ it follows from the way $\Lambda(f)$ is defined, 
that $\Lambda(\pi_2(f,A))\jdeq \Lambda(f)/A\jdeq
\Lambda(f)/A' \jdeq \Lambda(\pi_2(f,A'))$, and therefore we also get the equalities
\begin{equation*}
\pi_2(f,A)\jdeq \pi_2(\pi_2(f,A),())\jdeq \pi_2(\pi_2(f,A'),()) \jdeq \pi_2(f,A'),
\end{equation*}
so the desired equality holds.

@@

\begin{conj}
The functor from coherent split comprehension categories to CE-systems without
empty context is full and faithful.
\end{conj}

\begin{proof}
Let $\mathcal{E}$ be an CE-system. Define $\cat{E}$ to be the full subcategory
of $\cat{C}^\to$ whose objects are the morphisms of $\cat{F}$. Thus, for any
$\Gamma\in\cat{C}$, the category $p^{-1}(\Gamma)$ is the category of internal
morphisms the E-systems $\mathcal{E}/\Gamma$.

@@

\begin{thm}
The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{CEsys_{\circ}}$ to the
locally well-founded D-systems is full and faithful.
\end{thm}

@@

\begin{thm}
The image of $U$ is the subcategory of locally stratified CE-systems without
empty context. 
\end{thm}

\begin{proof}
For any $\Gamma\in\mathbb{D}$, there is an obvious stratification functor
$\Lambda(\Gamma)\to (\N,\geq)$, so $U(\mathbb{D})$ is a locally stratified
CE-system.

Now let $\mathbb{E}$ be a locally stratified CE-system without empty context.
For $\Gamma\in\cat{C}$, we define $\mathrm{Ty}(\Gamma)$ to be the set of
families of length one. Since any $f^\ast$ is stratified, it follows that
$\mathrm{Ty}$ is a presheaf on $\cat{C}$. The choice of pullback squares
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxext{\Delta}{f^\ast(A)} \arrow[r,"{\pi_2(f,A)}"] \arrow[d,fib] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,swap,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
is functorial by assumption, so we have defined a category with families. It
is straigtforward to verify that this category with families is locally well-founded,
and that its construction is inverse to $U$. 
\end{proof}
